---
title: "FFA Rasch Modelling"
format: html
author: Sebastian Sauer
toc: true
number-sections: true
---



```{r global-knitr-options, include=FALSE}
  knitr::opts_chunk$set(
  fig.pos = 'H',
  fig.asp = 0.618,
  fig.align='center',
  fig.width = 5,
  out.width = "100%",
  fig.cap = "", 
  fig.path = "chunk-img/",
  dpi = 300,
  # tidy = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  fig.show = "hold")
```



# Setup


```{r}
#| message=FALSE
library(easystats)  # convinient stats
library(here)  # path hell
library(eRm)  # Rasch models
library(mirt)  # more IRT
library(MASS)  # stats
library(lordif)  # DIF
library(tidyverse)  # data wrangling
library(psych)  # factor analysis
library(lavaan)  # CFAs
library(semPlot)  # SEM plots
```



## Import FFA data




```{r}
d_filename <- "Achtsamkeit_Daten_FFAEichung_rekodiert_2.sav"

d <- data_read(here("raw-data", d_filename))
```



## Median split

```{r}
d2 <- 
  d %>% 
  filter(Geschlecht != "divers") %>% 
  mutate(Geschlecht = as.character(Geschlecht)) %>% 
  mutate(PHQ_medsplit = ifelse(PHQ_Sum >= median(PHQ_Sum), 1, 0),
         Alter_medsplit = ifelse(Alter >= median(Alter), 1, 0))
```


Yes, only two cases of sex "divers".

## Check

Can we drop "diverse" sex without loosing much data?


```{r}
d %>% 
  count(Geschlecht)
```



# Prepare data


## Select FFA items

```{r}
ffa_items <-
  d2 %>% 
  dplyr::select(starts_with("FFA_"))
```


## Recode


```{r}
ffa_items2 <-
  ffa_items %>% 
  mutate(across(.cols = everything(),
                .fns = ~ case_when(. == "fast nie" ~ 0,
                                   . == "eher selten" ~ 1,
                                   . == "relativ oft" ~ 2,
                                   . == "fast immer" ~ 3)))
```



## Check

```{r}
ffa_items2 %>% 
 describe_distribution()
```




## FFA-13


```{r ffa13}
ffa13_items <-
  ffa_items2 %>% 
  dplyr::select(-FFA_13_rek)

dim(ffa13_items)
```



# IRT Models FFA13 - one factor model


## Rating Scale Model (RSM)


```{r}
#| eval: false
ffa_rsm1 <- RSM(ffa_items2, se = FALSE)
ffa_rsm1_ppar <- person.parameter(ffa_rsm1)
```


This model does not run, throws error due to singularity in matrix.




## Partial Credit Model (PCM)


### Estimate parameters

#### FFA-14: 14 items

```{r pcm1-ffa14}
ffa_pcm1 <- PCM(ffa_items2)
thresholds(ffa_pcm1)

ffa_pcm1_ppar <- person.parameter(ffa_pcm1)
ffa_pcm1_ppar

ffa_pcm1_itemfit <- eRm::itemfit(ffa_pcm1_ppar)
ffa_pcm1_itemfit
```

This model shows a bad fit.



#### FFA-13: 13 item

```{r pcm-13}
ffa_pcm2 <- PCM(ffa13_items)
ffa_pcm2_ppar <- person.parameter(ffa_pcm2)
ffa_pcm2_itemfit2 <- eRm::itemfit(ffa_pcm2_ppar)
ffa_pcm2_itemfit2
```






### Overall Goodness of Fit 

```{r}
#| eval: false
#gofIRT(ffa_pcm2_ppar)  # not implemented for polytomuous models yet
#item_info(ffa_pcm2)
```






### Andersen test

Significant results indicate that the exercise parameter differ significantly between the two groups.

```{r}
LRtest(ffa_pcm2)  # nicht so gut
LRtest(ffa_pcm2, d2$Evang)  # ok
LRtest(ffa_pcm2, d2$Theorie)  # nicht so gut

LRtest(ffa_pcm2, d2$PHQ_medsplit)  # nicht so gut
```





### Wald Test

```{r}
#| error: true
Waldtest(ffa_pcm2)
Waldtest(ffa_pcm2, d2$PHQ_medsplit)
```



### MLoef Test

```{r}
MLoef(ffa_pcm2)  # good
```


### Item information 

```{r}
#| error: true
plotINFO(ffa_pcm2, legpos = FALSE)
plotPImap(ffa_pcm2)
```



# Two factor model



## Devise factors

Items were assigned according to Sauer et al., 2013.


```{r}
presence_factor <- c(1, 2, 3, 5, 7)
acceptance_factor <- c(4, 6, 8, 9, 10, 11, 12, 14)
```



```{r ffa13-two-factors-tibbles}
ffa13_pres <-
  ffa_items2 %>% 
  dplyr::select(any_of(presence_factor))

ffa13_acc <-
  ffa_items2 %>% 
  dplyr::select(any_of(acceptance_factor))
```


## Check distributions


```{r}
ffa13_pres %>% 
  describe_distribution()
```


```{r}
ffa13_acc %>% 
  describe_distribution()
```


## Any categories not chosen?

Let's check if any category was not chosen at all, which may cause the model to fail [according to this source](https://stackoverflow.com/questions/67338560/zeroinfl-model-warning-message-in-sqrtdiagobjectvcov-nans-produced).



```{r}
ffa13_acc %>% 
  pivot_longer(everything(), names_to = "item", values_to = "category") %>% 
  count(item, category, sort = TRUE)
```

However, there appear to be no category with zero count.




# RSM


## Presence

### Parameter estimation

```{r 2f-rsm1-pres}
#| eval: TRUE
ffa_rsm1_pres <- RSM(ffa13_pres, se = TRUE)
ffa_rsm1_pres
ffa_rsm1_pres_ppar <- person.parameter(ffa_rsm1_pres)
```


### Itemfit

```{r}
eRm::itemfit(ffa_rsm1_pres_ppar)
```

Looks good.


### Andersen test

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.

```{r}
LRtest(ffa_rsm1_pres)  # sign.
LRtest(ffa_rsm1_pres, d2$Evang)  # sign
LRtest(ffa_rsm1_pres, d2$Theorie)  # sign

#PHQ_medsplit <- ifelse(d$PHQ_Sum >= median(d$PHQ_Sum), "+", "-")
LRtest(ffa_rsm1_pres, d2$PHQ_medsplit)  # sign

```





### Wald Test

The Wald Tests computes the difference in item difficulty normalized to their SE between two groups.

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.


```{r}
Waldtest(ffa_rsm1_pres)  # sign mostly
Waldtest(ffa_rsm1_pres, d2$PHQ_medsplit)  # signif mostly
```


### MLOEF Test

The M Loef tests checks whether the person parameter differ between items. 
According to the assumptions of the Rasch models, we expect invariance,
i.e., there should be not subset of items for which the person parameters differ from the rest of the items.


```{r}
MLoef(ffa_rsm1_pres)  # signif
```



### Item information 

```{r}
#| error: true
plotINFO(ffa_rsm1_pres, legpos = FALSE)
plotPImap(ffa_rsm1_pres)
```




## Acceptance

### Parameter estimation

This does NOT Run:

```{r 2f-rsm1-acc, message=FALSE}
#| eval: FALSE
ffa_rsm1_acc <- RSM(ffa13_acc, se = TRUE)
ffa_rsm1_acc
ffa_rsm1_acc_ppar <- person.parameter(ffa_rsm1_acc)
```

Oh no: NaNs have been produced. There must be some edge cases in the data.




### Itemfit

```{r}
#| eval: false
eRm::itemfit(ffa_rsm1_acc_ppar)
```





### Itemfit

```{r}
eRm::itemfit(ffa_rsm1_pres_ppar)
```



### Andersen test

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.

```{r}
#| error: true
LRtest(ffa_rsm1_acc)  # does not run
LRtest(ffa_rsm1_acc, d2$Evang)  # sign
LRtest(ffa_rsm1_acc, d2$Theorie)  # sign

#PHQ_medsplit <- ifelse(d$PHQ_Sum >= median(d$PHQ_Sum), "+", "-")
LRtest(ffa_rsm1_acc, d2$PHQ_medsplit)  # sign

```





### Wald Test

The Wald Tests computes the difference in item difficulty normalized to their SE between two groups.

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.


This function yields error (singular matrix).

```{r}
#| error: true
#Waldtest(ffa_rsm1_acc)  # sign mostly
#Waldtest(ffa_rsm1_acc, d2$PHQ_medsplit)  # signif mostly
```


### MLOEF Test

The M Loef tests checks whether the person parameter differ between items. 
According to the assumptions of the Rasch models, we expect invariance,
i.e., there should be not subset of items for which the person parameters differ from the rest of the items.


```{r}
#| error: true
MLoef(ffa_rsm1_acc)  # signif
```



### Item information 

```{r}
#| error: true
plotINFO(ffa_rsm1_acc, legpos = FALSE)
plotPImap(ffa_rsm1_acc)
```



## Conclusion

The RSM is not appropriate for the data for some reason that is not entirely clear.



# PCM


## Presence

### Parameter estimation

```{r 2f-pcm}
#| eval: true
ffa_pcm1_pres <- PCM(ffa13_pres, se = TRUE)
ffa_pcm1_pres
ffa_pcm1_pres_ppar <- person.parameter(ffa_pcm1_pres)
```


### Itemfit

```{r}
eRm::itemfit(ffa_pcm1_pres_ppar)
```

Quite good.



### Andersen test

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.

```{r}
LRtest(ffa_pcm1_pres)  # sign.
LRtest(ffa_pcm1_pres, d2$Evang)  # NOT sign
LRtest(ffa_pcm1_pres, d2$Theorie)  # sign

#PHQ_medsplit <- ifelse(d$PHQ_Sum >= median(d$PHQ_Sum), "+", "-")
LRtest(ffa_pcm1_pres, d2$PHQ_medsplit)  # sign

```





### Wald Test

The Wald Tests computes the difference in item difficulty normalized to their SE between two groups.

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.


```{r}
Waldtest(ffa_pcm1_pres)  #  mostly NOT signif
Waldtest(ffa_pcm1_pres, d2$PHQ_medsplit)  # mixed picture as to signif.
```






### MLOEF Test

The M Loef tests checks whether the person parameter differ between items. 
According to the assumptions of the Rasch models, we expect invariance,
i.e., there should be not subset of items for which the person parameters differ from the rest of the items.


```{r}
MLoef(ffa_pcm1_pres)  # signif
```



### Item information 

```{r}
#| error: true
plotINFO(ffa_pcm1_pres, legpos = FALSE)
plotPImap(ffa_pcm1_pres)
plotICC(ffa_pcm1_pres,mplot=TRUE,legpos=FALSE,ask=FALSE)
```






## Acceptance




### Parameter estimation

```{r 2f-pcm-acc}
#| eval: true
ffa_pcm1_acc <- PCM(ffa13_acc, se = TRUE)
ffa_pcm1_acc
ffa_pcm1_acc_ppar <- person.parameter(ffa_pcm1_acc)
```




### Itemfit

```{r}
#| eval: true
eRm::itemfit(ffa_pcm1_acc_ppar)
```

Quite good.







### Andersen test

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.

```{r}
#| error: true
LRtest(ffa_pcm1_acc)  # signif
LRtest(ffa_pcm1_acc, d2$Evang)  # NOT sign
LRtest(ffa_pcm1_acc, d2$Theorie)  # sign

#PHQ_medsplit <- ifelse(d$PHQ_Sum >= median(d$PHQ_Sum), "+", "-")
LRtest(ffa_pcm1_acc, d2$PHQ_medsplit)  # sign

```





### Wald Test

The Wald Tests computes the difference in item difficulty normalized to their SE between two groups.

A significant results indicates that the item parameters differ between groups, which indicates a violation of the Rasch model's assumption.


```{r}
Waldtest(ffa_pcm1_acc)  #  mostly not signif
Waldtest(ffa_pcm1_acc, d2$PHQ_medsplit)  #  NOT signif
```


### MLOEF Test

The M Loef tests checks whether the person parameter differ between items. 
According to the assumptions of the Rasch models, we expect invariance,
i.e., there should be not subset of items for which the person parameters differ from the rest of the items.


```{r}
MLoef(ffa_pcm1_acc)  # NOT signif
```



### Item information 

```{r}
#| error: true
plotINFO(ffa_pcm1_acc, legpos = FALSE)
plotPImap(ffa_pcm1_acc)
```



# Model comparison


## Presence


```{r}
anova(ffa_rsm1_pres, ffa_pcm1_pres)
```

```{r}
eRm::IC(ffa_rsm1_pres_ppar)
eRm::IC(ffa_pcm1_pres_ppar)
```

Small values are better.

That means the PCM models is to be preferred.



# DIF Analysis using ordinal logistic regression


## Presence 

### Gender

```{r}
ffa13_pres_dif_geschlecht <- lordif(ffa13_pres, group = d2$Geschlecht, pseudo.R2 = "McFadden", minCell = 5)
print(ffa13_pres_dif_geschlecht)
summary(ffa13_pres_dif_geschlecht)
plot(ffa13_pres_dif_geschlecht, labels = c("Frauen", "M채nner"))
```

Small DIF only, according to the R squeared value.



### Age

```{r}
ffa13_pres_dif_age <- lordif(ffa13_pres, group = d2$Alter_medsplit, pseudo.R2 = "McFadden", minCell = 5)
print(ffa13_pres_dif_age)
summary(ffa13_pres_dif_age)
plot(ffa13_pres_dif_age, labels = c("Frauen", "M채nner"))
```


Small DIF only, according to the R squeared value.





## Acceptance 

### Gender


Small DIF only, according to the R squeared value.



```{r}
ffa13_acc_dif_geschlecht <- lordif(ffa13_acc, group = d2$Geschlecht, pseudo.R2 = "McFadden", minCell = 5)
print(ffa13_acc_dif_geschlecht)
summary(ffa13_acc_dif_geschlecht)
plot(ffa13_acc_dif_geschlecht, labels = c("Frauen", "M채nner"))
```


Small DIF only, according to the R squared value.



### Age

```{r}
ffa13_acc_dif_age <- lordif(ffa13_pres, group = d2$Alter_medsplit, pseudo.R2 = "McFadden", minCell = 5)
print(ffa13_acc_dif_age)
summary(ffa13_acc_dif_age)
plot(ffa13_acc_dif_age, labels = c("Frauen", "M채nner"))
```


Small DIF only, according to the R squeared value.



# Dimensionality

## EFA

## Determine number of factors


2 factors, oblimin rotation, ML factor scores:


```{r fa1}
ffa_fa <- psych::fa(ffa13_items, nfactors = 2, rotate = "oblimin", fm = "ml")
print(ffa_fa)
print(ffa_fa$loadings, cutoff = 0.2)
```



Try different number of factors:


```{r}
ffa_div_fa <- nfactors(ffa13_items, n = 5, fm = "ml")
ffa_div_fa
```

This analyses both supports the 1- and the 2-F solution.


```{r}
scree(ffa13_items, pc = FALSE)
```
Screeplot (in combination with Kaiser criterion) suggests a unidimensional solution.

```{r}
vss(ffa13_items, fm = "ml", n = 3, plot = FALSE)
```

VSS seems to suggest a 1 factor solution too.

MAP favors the 1 factor solution too.

### Goodness of fit

```{r}
summary(ffa_fa)
```


## CFA

### 2F model

```{r}
ffa_mod <- '
pres =~ FFA_1 + FFA_2 + FFA_3 + FFA_5 + FFA_7
acc =~ FFA_4 + FFA_6 + FFA_8 + FFA_9 + FFA_10 + FFA_11 + FFA_12 + FFA_14'

ffa_cfa <- lavaan::cfa(ffa_mod, data = ffa_items2, ordered = TRUE)
#ffa_cfa <- lavaan::cfa(ffa_mod, data = ffa_items2)

ffa_cfa
```

```{r}
semPaths(ffa_cfa, what = "est", intercepts = FALSE, 
         rotation = 4, edge.color = 1, fade = FALSE, 
         edge.label.cex = .5, edge.width = .3)
```


```{r}
summary(ffa_cfa, standardized = TRUE, fit.measures = TRUE)
```


### 1F model

```{r}
ffa_mod_1F <- 'm =~ FFA_1 + FFA_2 + FFA_3 + FFA_5 + FFA_7 + FFA_4 + FFA_6 + FFA_8 + FFA_9 + FFA_10 + FFA_11 + FFA_12 + FFA_14'

ffa_cfa_1F <- lavaan::cfa(ffa_mod_1F, data = ffa_items2, ordered = TRUE)
#ffa_cfa <- lavaan::cfa(ffa_mod, data = ffa_items2)

ffa_cfa_1F
```




```{r}
semPaths(ffa_cfa_1F, what = "est", intercepts = FALSE, 
         rotation = 4, edge.color = 1, fade = FALSE, 
         edge.label.cex = .5, edge.width = .3)
```


```{r}
summary(ffa_cfa_1F, standardized = TRUE, fit.measures = TRUE)
```




### Conclusion


The CFA seems to speaks in favor of the 2F solution.
